---
title: "457 tut1"
author: '"Yukun Gao"'
date: "2022/1/25"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## (a)

```{r}
# set seed so generate same output
set.seed(011) 
 #signal
s_t1 <- c(rep(0,100), 10*exp(-(1:100)/20)*cos(2*pi*1:100/4))
#signal plus white noise with mean 0 and sd of 1, x_t= s_t+w_t
x_t1 <- s_t1+ rnorm(200)
# plot time series
plot.ts(x_t1, main="Signal plus noise Model in (a)",
        col="green", ylab=expression("xt for (a)"))
```
The plot answers (c)


## (b)

```{r}
# set seed so generate same output
set.seed(011)
#signal
s_t2 <- c(rep(0,100), 10*exp(-(1:100)/200)*cos(2*pi*1:100/4))
#signal plus white noise with mean 0 and sd of 1, x_t= s_t+w_t
x_t2 <- s_t2+ rnorm(200) 
# plot time series
plot.ts(x_t2, main="Signal plus noise Model in (b)",
        col="red",ylab=expression("xt for (b)"))
```
The plot answers (c)


## (c)

By looking at the two graph, we notice that the first half of the graph looks pretty similar. The signal $s_t$ begins at t=101, we see both model stay at center around 0 and show decrease in variance as t increases. However, the graph for Model in (a) shows faster decrease in variance than Model in (b).

## (d)

```{r}
#generate signal modulator (i): exp{-t/20} 
i<- exp(-(1:100)/20)
#generate signal modulator (ii): exp{-t/200} 
ii<- exp(-(1:100)/200)

# plot signal modulator (i)
plot.ts(i, col="green", ylab="signal modulator")
#plot signal modulator (ii)
lines(ii, col="red")
#add legend
legend("topright", legend = c("(i)","(ii)"),
       cex=0.5,col=c("green","red"), lty=1)
```


By comparing signal modulator (i) and (ii), we clearly see that signal (i) decrease faster than (ii).


## (e)

$E(x_t) =E(s_t+w_t)= E(s_t) +E(w_t)= E(s_t)=\mu_x{t}$ which is constant, since $E(w_t)=0$ So we can just plot $s_t$ for the two models.


```{r}
#plot mean function for model in (a)
plot.ts(s_t1 ,ylab=expression(mu[a]),
        main="Mean of Model in (a)",col="green")
#plot mean function for model in (b)
plot.ts(s_t2 ,ylab=expression(mu[b]),
        main="Mean of Model in (b)",col="red")
```

