---
title: "Tut457"
author: '"Yukun Gao"'
date: "2022/3/10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


$x_t-1.5x_{t-1}+0.75 x_{t-2} = w_t$ where $\phi_1=1.5$ and $\phi_2=-0.75$

We could write $\phi(z)= 1-1.5z+0.75z^2$ , set $\phi(z)=0$. We get $z_1 = 1+\frac{\sqrt{3}}{3}i=1+0.57735i$ and  $z_2= 1-\frac{\sqrt{3}}{3}i=1-0.57735i$.
According to 3.10, $\gamma(h)=\phi_1\gamma(h-1)+\phi_2\gamma(h-2)$, for h=1,2,.. for causal AR(2) process. Both roots are outside the unit circle, so the model is causal. We can divide by $\gamma(0)$ to get the difference equation for the ACF.
$\rho(h)=\phi_1\rho(h-1)-\phi_2\rho(h-2)$ for h=1,2,.... and we know $\rho(0)=1$
and $\rho(1)=\rho(-1)=\frac{\phi_1}{1-\phi_2}$


We know if  $z_1$ and $z_2$ are in conjugate pairs, $\rho(h)=a|z_1|^{-h}cos(h\theta+b)$. Then if $h=0$, $\rho(0)=acos(b)=1$ and $a= \frac{1}{cos(b)}$. If $h=1$, $\rho(1)=a*\frac{1}{|z_1|}cos(\theta+b)= \frac{\phi_1}{1-\phi_2}$. Dividing these two equations, we get $\frac{cos(\theta +b)}{cos(b)}= \frac{cos(\theta)cos(b)}{cos(b)}-\frac{sinc(\theta)sin(b)}{cos(b)}=cos(\theta)-sin(\theta)tan(b)=|z_1|*\frac{\phi_1}{1-\phi_2}$, which means $b= tan^{-1} \{ \frac{cos(\theta)-|z_1|*\frac{\phi_1}{1-\phi_2}}{sin(\theta)} \}$. We will use the formula for $a$ and $b$.

```{r}
#find roots of polynomial
x=c(1,-1.5,0.75)
root_poly<-polyroot(x)
#the complex part is 0.57735, find theta
theta<- atan(0.57735/1)
#calculate abs(z1)=1.154701
z1norm<- abs(root_poly[1])

##calculate b
b<- atan((cos(theta)-z1norm*1.5/(1+0.75))/sin(theta))
## compute a
a<-1/cos(b)

#calculate the value of rho(h)
h=c(0:10)
rhoh<- a*z1norm^(-h)*cos(h*theta+b)
rhoh
```

```{r}
#plot the values using
plot(rhoh, type = "h", xlab="lag",ylab="ACF" ,main="ACF plot for AR(2)")
abline(h=0)
```
```{r}
#generate ARMAacf to check answer
new_acf <-ARMAacf(ar=c(1.5,-0.75),ma=0, lag.max=10)
new_acf
plot(new_acf, type = "h", xlab="lag",ylab="ACF" ,main="ACF plot using ARMAacf")
abline(h=0)
```

we see that the plot of  ARMAacf correspond to our answers. 
