---
title: "tut7"
author: '"Yukun Gao"'
date: "2022/3/30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
#load the library
library(astsa)
#Loads specified data sets
data(globtemp)
```

## (a)
```{r}
#plot the data
plot(globtemp, xlab = "Year", ylab = "Global Temperature",
col = "blue", main = "Global Temperature")
```
```{r}
#plot ACF
plot1 <- acf2(globtemp, max.lag= 60,main = "ACF and PACF Plots")
```
The sample ACF plot shows a slow decay to zero we might need to defference the data



```{r}
#difference the data
firstdiff <- diff(globtemp)

#plot the new data
plot(firstdiff, xlab = "Year", ylab = "Global Temperature",
main = "Global Temperature after Difference")
```
The first difference of Global Temperature data seems to be a stable process


```{r}
#Plot the ACF and PACF of the "first difference" data
plot2 <-acf2(firstdiff, main = "ACF and PACF Plots", max.lag=60)
```

PACF approcimately cutoff at lag 3,  ACF approximately cutting off at lag 2. We consider there might be two models: 1. ARIMA(3, 1, 0) and ARIMA(0, 1, 2), and we will compare them. 

## (b)
```{r}
#ARIMA(3, 1, 0) model
sarima(globtemp, 3, 1, 0,details=FALSE)
```

For the proposed ARIMA(3, 1, 0) model, except fort the constant term, the p-values for all other parameter estimates are smaller than 0.05, which is significant.


```{r}
#ARIMA(0, 1, 2) model
sarima(globtemp, 0, 1, 2,details=FALSE)
```
For the proposed ARIMA(0, 1, 2) model, the p-values for all parameter estimates are smaller than 0.05, which is significant.

## (c)
```{r}
#analyse ARIMA(3, 1, 0) model
Model1 <-sarima(globtemp, 3, 1, 0)
#analyse ARIMA(0, 1, 2) model
Model2<-  sarima(globtemp, 0, 1, 2)
```


There are no obvious patterns of the standardized residuals plot for both model.


There is a spike at lag 18 for both modal in the ACF of Residuals plots, but it is not very significant at the 5% level which does not violate the randomness assumption. 

The Normal Q-Q plots show that the normality assumption is valid, for both model, although there are some outliers. 

The p-values of Ljung-Box statistics are all above the significant level for both models, which we cannot reject the null hypothesis that the residuals are
independent.

The two models's diagnostic from the plots are similar, we cannot select a better model. However, by comparing the AIC/AICc/BIC for two models, we will propose for ARIMA(0,1,2) to be the better model. Since the AIC/AICc/BIC are all smaller for ARIMA(0,1,2) model. 

## (d)
```{r}
#Prediction for 10 periods
prediction <- sarima.for(globtemp, n.ahead= 10, 0, 1, 2)
# 5% Prediction interval
upper95 <- prediction$pred + qnorm(0.95) * prediction$se
lower95 <-prediction$pred - qnorm(0.95) * prediction$se


data.frame("Prediction" = prediction$pred,
"Lower Bound" = lower95, "Upper Bound" = upper95)

```


