---
title: "tut8-1"
author: '"Yukun Gao"'
date: "2022/4/7"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## a)

```{r}
library(astsa)
library(MASS)
data(salt)
```

```{r}
##plot time series
plot.ts(salt)
# The spectral analysis
analysis_spec = mvspec(salt, log = "no")
```
```{r}
# Find the first three dominant frenquencies
dom_freq <-analysis_spec$details[order(analysis_spec$details[,3],
                                       decreasing = TRUE),]
#show the  three dominant frequencies 
dom_freq[1,] 
dom_freq[2,] 
dom_freq[3,]
```
The dominant frequencies are: 0.0625, 0.0469, 0.1094       


## b)

```{r}
#Cycle is 1/ frequency
1/dom_freq[1,1] 
1/dom_freq[2,1] 
1/dom_freq[3,1]
```
As we can see, the cycles occurs at 16, 21.32196, 9.140768

## c)

```{r}
##Obtain 90% CIs for the three dominant frequencies
freq_ci_1 = 2*dom_freq[1,3]/qchisq(0.05,2)
freq_ci_2 = 2*dom_freq[1,3]/qchisq(0.95,2)
freq_ci_3 = 2*dom_freq[2,3]/qchisq(0.05,2)
freq_ci_4 = 2*dom_freq[2,3]/qchisq(0.95,2)
freq_ci_5 = 2*dom_freq[3,3]/qchisq(0.05,2)
freq_ci_6 = 2*dom_freq[3,3]/qchisq(0.95,2)

##Create data frame for the CI of three dominant frequencies and show them out
data.frame(Dom_freq=c(dom_freq[1,1],dom_freq[2,1],dom_freq[3,1]), Spec=c(dom_freq[1,3],
                                                                         dom_freq[2,3], dom_freq[3,3]), LowerBound=c(freq_ci_2,freq_ci_4,freq_ci_6),
UpperBound=c(freq_ci_1,freq_ci_3,freq_ci_5))

```

Interpretation:

1. For the first dominant frequency, we see the periodogram ordinate is 60.67, which lies within the  confidence intervals of the second and third peak. It is not significant.

2. For the second dominant frequency, we see the periodogram ordinate is 33.4859, which lies within the confidence intervals of the first  and third peak. It is not significant.

3. For the third dominant frequency, we see the periodogram ordinate is 23.6903, which lies within the confidence intervals of the second  and third peak. It is not significant.
